Proposed Solution

### 2017-04-18 OOPSLA17/secs/intro.tex
proper outline
autocommit

Configuration errors (also known as misconfigurations) have become
one of the major causes of system failures, resulting in security vulnerabilities,
application outages, and incorrect program executions~\cite{xu15systems, xu13do, xu15hey}. 
In one high profile case,  
Facebook and Instagram became inaccessible~\cite{mashableNews}
and a Facebook spokeswoman reported that 
this was caused by a change to the site's configuration systems.
In another case, a system configuration change caused an outage of AT\&T's 911 service for five hours.
During this time, 12,600 unique callers were not able to reach the 911 emergency line~\cite{att-outage}.
These critical system failures are not rare either -
a software system failures study~\cite{yin11anempirical},
reports that about 31\% of system failures were caused by 
configuration errors.
This is even higher than the percentage of failures resulting from program bugs (20\%).

The systems research community has recognized this as an important
problem~\cite{xu16early} and many efforts have been proposed to
check, troubleshoot, and diagnose configuration 
errors~\cite{attariyan10automating,
su07autobash, whitaker04configuration,xu2016early}.
However, these tools rely on analyzing the source code of 
the target systems or need to run the system
multiple times to understand the source of the errors.
The support for a static analysis style verification for configuration files
is still not on a level of
automated verification tools used for regular program 
verification~\cite{Leino10Dafny, PiskacWZ14, BobotFMP15} that can
preemptively detect errors.

In this paper, we propose a framework for automated verification of configuration files: 
we derive specifications for a correct configuration file by processing a training set of configuration files 
and proactively report potential errors in the configuration files of interest.
In this framework, the example configurations from which we derive the specifications do not need to be labeled as correct or incorrect.

There are two main obstacles to directly applying existing automatic program
verification techniques to verification of configuration files:
1) the lack of a specification describing properties of configuration files;
2) the flat structure of configuration files -- they
are mainly a sequence of entries assigning some value to system
variables (called {\emph {keywords}}). 
The language in which configuration files are written does 
not always adhere to a specific grammar or syntax. In particular, the
entries in configuration files are untyped. 
Moreover, there are surprisingly few rules specifying constraints on entries, 
and there is no explicit structure policy for the entries.

\app overcomes the above obstacles by first automatically inferring a
specification for configuration files. It is unrealistic to expect the 
users to write an entire specification for configuration files on their own. 
This process can easily lead to incomplete or even contradictory 
specifications. Instead, we learn specification from 
a large training set of 
configuration files~\cite{configdataset}. The first step is to 
translate this training set into a more structured typed representation.
We then apply the learning process and we learn an abundant set of rules 
specifying various properties that hold on the given training set. The rules, in general, 
specify which properties keywords in configuration files need to satisfy.
The learning process is language-agnostic and works for any kind 
of configuration files,  but all of the files in the training set need to be of 
the same kind (such as MySQL or HTTPD configuration files).
This learning process is a way to derive a specification for configuration files. 
%It is hard to talk here about a complete specification, but it is still
%a set of formal rules describing the properties that configuration files need to satisfy.

Having a specification, \app can then efficiently check 
the correctness of the configuration files of interest and detect 
potential errors. Errors are reported if the configuration file does not 
adhere to the derived specification.

The errors \app detects may cause total system failures, but can also be more insidious, for example slowing down the system only when the server load increases beyond a certain threshold.
Since these runtime errors may only be triggered after some time in a deployment environment, the standard debugging 
techniques~\cite{Zeller:2005:WPF:1077048} of starting a system multiple times with different configuration settings will not help detect these misconfigurations.

\markk{xu16early should go here somehow}
There have been several prior efforts that attempt to proactively verify 
configuration files~\cite{santolucitoCAV, xu16early,
zhang14encore, huang15confvalley}.
However, state-of-the-art efforts have limitations that are impractical to meet in practice.
In general, these efforts fall into two categories with respect to their limitations.\begin{itemize}
\item On the one hand there are tools that can detect sophisticated 
configuration errors, \eg, ConfigC~\cite{santolucitoCAV}. 
However, these tools heavily rely on datasets containing 100\% 
correct configuration files to extract configuration rules.
Existing investigation studies~\cite{wang04automatic, yin11anempirical}
have demonstrated determining or obtaining 100\% correct configuration
files to drive rules is almost impossible in reality. 
Without 100\% correct datasets, these tools do not work.
\item On the other hand there are tools, 
\eg, EnCore~\cite{zhang14encore} and
ConfValley~\cite{huang15confvalley}, that can only 
detect rather simplistic configuration errors (\eg, value range errors 
and simple integer correlation errors), but cannot detect
more complex configuration errors, such as ordering errors or 
Nevertheless, these types of errors cannot be detected by above existing efforts.
\end{itemize}

Our learning algorithm surmounts both of these shortcomings present in 
existing tools.  \app implements a learning
algorithm inspired by {\em association rule
learning}~\cite{agrawal1993mining}. It analyzes a large training set of configuration files~\cite{configdataset}. Those are real-world reported misconfigurations. A file in the training set might 
contain several different errors and these errors only appear 
in a small percentage of files. We first translate those file into an 
intermediary typed language. With with every type, we associate 
a set of very general interfaces. The user does not
need to provide them - they are internally associated with the types. The 
learner then instantiates those interfaces with the keywords appearing in 
the training set. However, since files might also contain errors we take 
this into account when learning correct rules. All learned rules are 
annotated with the probability of correctness.  
To ensure the learned rules are correct enough,
\app employs a graph analysis to refine the set learned rules.
\app analyzes the rule graph to rank the importance and relevance of the learned rules.

>>>>

%what is the problem
Configuration errors (also known as misconfigurations) have become
one of the major causes of system failures, resulting in security vulnerabilities,
application outages, and incorrect program executions~\cite{xu15systems, xu13do, xu15hey}. 
In one high profile case,  
Facebook and Instagram became inaccessible~\cite{mashableNews}
and a Facebook spokeswoman reported that 
this was caused by a change to the site's configuration systems.
In another case, a system configuration change caused an outage of AT\&T's 911 service for five hours.
During this time, 12,600 unique callers were not able to reach the 911 emergency line~\cite{att-outage}.
These critical system failures are not rare -
a software system failures study~\cite{yin11anempirical},
reports that about 31\% of system failures were caused by 
configuration errors.
This is even higher than the percentage of failures resulting from program bugs (20\%).

%what has been done
The systems research community has recognized this as an important
problem and many efforts have been proposed to
check, troubleshoot, and diagnose configuration 
errors~\cite{attariyan10automating,
su07autobash, whitaker04configuration,xu16early}.
However, these tools rely on analyzing the source code of 
the target systems or need to run the system
multiple times to understand the source of the errors.
The support for a static analysis style verification for configuration files
is still not on the level of
automated verification tools used for regular program 
verification~\cite{Leino10Dafny, PiskacWZ14, BobotFMP15} that can
preemptively detect errors.

%why is that insufficent
%\markk{xu16early should go here somehow}
There have been several prior efforts that attempt to proactively verify 
configuration files~\cite{santolucitoCAV, xu16early,
zhang14encore, huang15confvalley}.
However, state-of-the-art efforts have limitations that are impractical to meet in practice.
In general, these efforts fall into two categories with respect to their limitations.
\begin{itemize}
\item On the one hand there are tools that can detect sophisticated 
configuration errors, \eg, ConfigC~\cite{santolucitoCAV}. 
However, these tools heavily rely on datasets containing 100\% 
correct configuration files to extract configuration rules.
Existing investigation studies~\cite{wang04automatic, yin11anempirical}
have demonstrated determining or obtaining 100\% correct configuration
files to drive rules is almost impossible in reality. 
Without 100\% correct datasets, these tools do not work.
\item On the other hand there are tools, 
\eg, EnCore~\cite{zhang14encore} and
ConfValley~\cite{huang15confvalley}, that can only 
detect rather simplistic configuration errors (\eg, value range errors 
and simple integer correlation errors), but cannot detect
more complex configuration errors, such as ordering errors, fine-grain correlations, missing entry errors.
\end{itemize}

%what do we do
In this paper, we propose a framework for automated verification of configuration files: 
We derive specifications for a correct configuration file by processing a training set of configuration files that may contain some errors.
The errors that we learn are detailed and complex enough proactively report a wide range of potential misconfigurations.
In this framework, the example configurations from which we derive the specifications do not need to be labeled as correct or incorrect.

% how did we do it
There are two main obstacles to directly applying existing automatic program
verification techniques to configuration files.
First, a lack of specifications on the properties of configuration files makes the verification task poorly defined.
There are surprisingly few rules specifying constraints on entries, even written in plain English.
Since asking users to write an entire specification for configuration files is impractical and error-prone, we instead take a specification generation approach.
Second, the flat structure of configuration files is only a sequence of entries assigning some value to system
variables (called {\emph {keywords}}) and provides little structural information.
In particular, the keywords in configuration files are often untyped, a useful property to leverage in program verification.

In order to learn a specification from a large training set of configuration files,
the first step is to 
translate this training set into a more structured typed representation.
We then apply the learning process and we learn an abundant set of rules 
specifying various properties that hold on the given training set. The rules, in general, 
specify which properties keywords in configuration files need to satisfy.
The learning process is language-agnostic and works for any kind 
of configuration files, though all of the files in the training set need to be of 
the same language (such as MySQL or HTTPD configuration files).
This learning process is a way to derive a specification for configuration files. 
With the learned specification, \app can then efficiently check 
the correctness of the configuration files of interest and detect 
potential errors. Errors are reported if the configuration file does not 
adhere to the derived specification.

The errors \app detects may cause total system failures, but can also be more insidious, for example slowing down the system only when the server load increases beyond a certain threshold.
Since these runtime errors may only be triggered after some time in a deployment environment, the standard debugging 
technique~\cite{Zeller:2005:WPF:1077048} of starting a system multiple times with different configuration settings will not help detect these misconfigurations.

%did what we did work?
Our learning algorithm surmounts both of these shortcomings present in 
existing tools.  \app implements a learning
algorithm inspired by {\em association rule
learning}~\cite{agrawal1993mining}. It analyzes a large training set of configuration files~\cite{configdataset}. Those are real-world reported misconfigurations. 
A file in the training set might contain several different errors and these errors only appear 
in a small percentage of files. 
We first translate those file into an intermediary typed language. 
With with every type, we associate a set of very general interfaces. 
The user does not need to provide them - they are internally associated with the types. 
The learner then instantiates those interfaces with the keywords appearing in 
the training set. However, since files might also contain errors we take 
this into account when learning correct rules. All learned rules are 
annotated with the probability of correctness.  
To ensure the learned rules are correct enough,
\app employs a graph analysis to refine the set learned rules.
\app analyzes the rule graph to rank the importance and relevance of the learned rules.
